The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2017-08-29 JStatSoft_forestinventory_master-concordance.tex
update Andreas 29.08.2017(1)

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 123 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
139 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
1 77 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
175 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
192}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
194 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
1 46 1 1 2 11 0 1 2 35 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1513:%
206 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1522:%
1 21 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1544:%
216 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1560:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1818:%
233 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1822:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1827:%
238 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 123 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 131:%
1 95 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 227:%
139 34 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 262:%
1 77 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 340:%
175 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 356:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 887:%
192}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 888:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1407:%
194 10 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1418:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1483:%
206 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1492:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 21 1 1 8 18 1 1 2 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1698:%
216 15 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1714:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 36 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1972:%
233 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1976:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1981:%
238 11 1}

### 2017-08-29 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 29.08.2017(1)

% NOTE:
% - Provide all confidence equations for the estimators: onephase, twophase, threephase
% - Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
% - ... and asymptotic of estimators
% - Interpretation of Confidence Intervals
%

Whereas the estimated variance of a point estimate provides a measure of estimation uncertainty, its interpretation is rather diffult to use by practitioners. Converting the estimated variance into a 95\% \textit{confidence interval} allows for a much more intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate. The correct interpretation of a confidence interval is not that there is a 95\% probability that it contains the true value.  In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sampling design. Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\% confidence intervals from each sample, then 95\% of them are expected to contain the true population parameter. The confidence level $1-\alpha$ (e.g. 95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown  population parameter and is therefore often also referred to as \textit{coverage rate}.

Based on the central limit theorem, it can be assumed that under hypothetical repeated sampling, the point estimates will asymptotically follow a normal distribution. \citet{mandallaz2013a} recommended to better base the confidence intervals on the Student's \textit{t} distribution. The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as \\

@@

where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level. Note that for three-phase sampling, $p$ constitutes the number of parameters used in the \textit{full} regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots).

In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function. The 95\% confidence intervall of the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) can for example be applied by:

@@

%  NOTE:
% - note that the CIs can be used for hypothesis testing, i.e. it can give us an idea wether
%   an estimate is statistically different from a given value
% - 
%
%

>>>>

% NOTE:
% - Provide all confidence equations for the estimators: onephase, twophase, threephase
% - !!! Give recommendations for n-threshold of unbiased estimators based on confid.level (dfs)
% - ... and asymptotics of estimators
% - Interpretation of Confidence Intervals
%

Converting the estimated variance into a 95\% \textit{confidence interval} (CI) allows for an intuitive and practice-oriented interpretation of the uncertainties associated to a point estimate. The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value.  In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sampling design. Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples, and then calculate 95\%  CIs from each sample, then 95\% of them are expected to contain the true population parameter. The confidence level $1-\alpha$ (e.g. 95\%) is thus the frequency or proportion of possible confidence intervals that contain the unknown population parameter and is therefore often also referred to as \textit{coverage rate}. For a user, the CI is a measure of precision that can also be used for hypothesis testing, i.e. providing evidence that an estimate is statistically different from a given value if that value lies outside the CI boundaries.

Based on the central limit theorem, it can be assumed that under hypothetical repeated sampling, the point estimates will asymptotically follow a normal distribution. \citet{mandallaz2013a} recommends to better base the confidence intervals on the Student's \textit{t} distribution. The confidence intervals for one-phase as well as two-phase and three-phase global and small area point estimates are thus defined as

@@

where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, and $1-\alpha$ is the confidence level. Note that for three-phase sampling, $p$ constitutes the number of parameters used in the \textit{full} regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). In the package, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic function \code{confint()}, which requires an estimation object created by the \code{onephase()}, \code{twophase()} or \code{threephase()} function. The 95\% confidence intervals for the small area estimates by the \textit{extpsynth} estimator (section \ref{sec:twophase_sae}) for example can  be applied by:

@@

### 2017-08-29 rnw/Main structure and modules of the Package.Rnw
update Andreas 29.08.2017(1)

The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories. However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest). As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e. the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable. Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e. the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process. The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism. A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.

>>>>

The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories. However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest). As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e. the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable. Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e. the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process. The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism. A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2013a}.

### 2017-08-29 rnw/Special Cases and Scenarios.Rnw
update Andreas 29.08.2017(1)

% include:
% - categorical variables (zberg) and interaction terms (not so clear to everyone)
% - !!! cluster sampling and differences between (p)small and extended (p)small if cluster is not completely inside small area
% - !!! violation of nested sample design + error handling / data adjustments / printed warnings
% - !!! missing factor levels in s2-sample that occur in s1 / s0 sample + error warning
% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
%

\subsection{Small Area Estimation under Cluster Sampling}

>>>>

% include:
% - categorical variables (zberg) and interaction terms (not so clear to everyone)
% - [mention that especially in the case of postratification, the g-weight variance is the preferred type because ...]
% - !!! cluster sampling and differences between (p)small and extended (p)small if cluster is not completely inside small area
% - !!! violation of nested sample design + error handling / data adjustments / printed warnings
% - !!! missing factor levels in s2-sample that occur in s1 / s0 sample + error warning
%

% --------------------------------------------------------- %
\subsection{Poststratification}

A special case of multi-phase regression estimators is \textit{poststratification}, which can further be divided into the case of \textit{multi-phase sampling for stratification} and \textit{multi-phase sampling for regression within strata}. Both imply the use of one or more categorical variables in the regression model(s), leading to classical ANOVA and ANCOVA models.

To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} dataset into 3 classes:

\begin{small}
<<>>=
grisons$stage<- as.factor(kmeans(grisons$mean, centers = 3)$cluster)
@
\end{small}

\textit{Two-phase sampling for stratification} is now applied if we only use categorical variables in the model, in this case the factor variable \code{stage} (i.e. we have an ANOVA model). It basically means that the model predictions are the means of terrestrial response values within each development stage (within-strata means).

\begin{small}
<<eval=FALSE>>=
# two-phase sampling for stratification
twophase(formula=tvol ~ stage,
         data=grisons,
         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
         boundary_weights = "boundary_weights")
@
\end{small}

\textit{Two-phase sampling for regression within strata} implies the combination of continuous and categorical variables within the model (i.e. we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators.

\begin{small}
<<eval=FALSE>>=
# two-phase sampling for regression within strata:
twophase(formula=tvol ~ mean + stddev + max + q75 + stage,
         data=grisons,
         phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
         boundary_weights = "boundary_weights")
@
\end{small}

The overall variance of estimators under poststratification will mainly depend on the between and within-strata variation of the model residuals and can therefore be reduced by minimizing the within strata residual square sum. In case of poststratification, the g-weight variance estimator has the advantage that the strata weights are estimated from the large sample, whereas the external variance calculates them solely based on the terrestrial sample $s_2$.

% --------------------------------------------------------- %
\subsection{Small Area Estimation under Cluster Sampling}

% NOTE:
% 
%
%

### 2017-08-29 rnw/Twophase_Estimators and their Application.Rnw
test on new branch
test2 newbranch
update Andreas 29.08.2017(1)

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

>>>>

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly smaller than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.
